Compare pseqs

### 2019-10-04 comparepseqs.Rmd
Rewrote

    if (!file.exists(paste0("rds-comparepseqs/alpha-", key , ".rds"))) {
        missing$alpha = TRUE
    } else {
        alphadiversity <- readRDS(file = paste0("rds-comparepseqs/alpha-", key , ".rds"))
    }

    if (!file.exists(paste0("rds-comparepseqs/bray-", key , ".rds"))) {
        missing$bray = TRUE
    } else {
        bray.dist.matrix <- readRDS(file = paste0("rds-comparepseqs/bray-", key , ".rds"))
    }

    if (!file.exists(paste0("rds-comparepseqs/beta-", key , ".rds"))) {
        missing$beta = TRUE
    } else {
        betadiversity <- readRDS(file = paste0("rds-comparepseqs/beta-", key , ".rds"))
    }

    if (length(missing) > 0) {

        pseq <- import_filter_data(rds[[key]],
                                   included = c("MEN", "BL_AGE", "SYSTM", "DIASM", "BMI",
                                                "CURR_SMOKE", "BL_USE_RX_C09",
                                                "Q57X", "BL_USE_RX_C03",
                                                "BL_USE_RX_C07", "BL_USE_RX_C08"))
        dset <- meta.merge.alphadiversity(pseq) %>%
            mutate(diversities_shannon = as.numeric(scale(diversities_shannon)))

        if ("alphadistribution" %in% names(missing)) {
            message("Missing alpha distribution")
            alphadistribution <- dset %>%
                mutate(!!key := diversities_shannon) %>%
                select(Sample_ID, key)
            saveRDS(alphadistribution,
                    file = paste0("rds-comparepseqs/alphadistribution-", key , ".rds"))
        }

        if ("alpha" %in% names(missing)) {
            message("Missing alpha")
            alphadiversity <- list(min =  calculateglm(dset, modelstring = fo.alpha.min),
                                   full = calculateglm(dset, modelstring = fo.alpha.full))
            saveRDS(alphadiversity, file = paste0("rds-comparepseqs/alpha-", key , ".rds"))
        }

        if ("bray" %in% names(missing)) {
            message("Missing Bray")
            bray.dist.matrix <- calculate.beta.matrix(pseq)
            saveRDS(bray.dist.matrix, file = paste0("rds-comparepseqs/bray-", key , ".rds"))
        }

        if ("beta" %in% names(missing)) {
            message("Missing beta")
            betadiversity <- list(min = calculateadonis(dset = dset,
                                                        matrix = bray.dist.matrix,
                                                        covariates = covariates.min,
                                                        npermutations = 1000),
                                  full = calculateadonis(dset = dset,
                                                         matrix = bray.dist.matrix,
                                                         covariates = covariates.full,
                                                         npermutations = 1000))
            saveRDS(betadiversity, file = paste0("rds-comparepseqs/beta-", key , ".rds"))
        }
    }

    ret[[key]] = list("alphadiversity" = alphadiversity,
                      "betadiversity" = betadiversity,
                      "alphadistribution" = alphadistribution)
}
```

```{r quit subruns}
if (exists("subset.run")) {
    quit()
}
```

# Diversities for BP variables

```{r plot alpha-beta}
plotlist <- list()

for (key in names(rds)) {
    diversities <- mergediversities(ret[[key]][["alphadiversity"]],
                                    ret[[key]][["betadiversity"]])

    alpha.diversity.min <- plot.alpha(diversities$min, alphabreaks = c(-0.15, -0.10, -0.05, -0.01, 0))
    alpha.diversity.full <- plot.alpha(diversities$full, alphabreaks = c(-0.15, -0.10, -0.05, -0.01, -0))

    beta.diversity.min <- plot.beta(diversities$min)
    beta.diversity.full <- plot.beta(diversities$full)

    gs4 <- list(alpha.diversity.min$yaxis,
                alpha.diversity.min$plot,
                beta.diversity.min$plot,
                alpha.diversity.min$legend,
                alpha.diversity.min$xaxis,
                beta.diversity.min$xaxis,
                beta.diversity.min$xlab,
                text_grob(sprintf("A. %s (%s)", key, paste(covariates.min, collapse = ", ")),
                          size = 16,
                          just = "left"),
                alpha.diversity.full$yaxis,
                alpha.diversity.full$plot,
                beta.diversity.full$plot,  
                alpha.diversity.full$xaxis,
                beta.diversity.full$xaxis,
                beta.diversity.full$xlab,
                text_grob(sprintf("B. %s (%s)", key, paste(covariates.full %>%
                                                           gsub("BL_USE_", "", .), collapse = ", ")),
                          size = 16, just = "left"))

    g4 <- arrangeGrob(grobs = gs4,
                      layout_matrix = rbind(
                          c(8,    NA,    NA, NA,   NA, NA, NA),
                          c(NA, 1,    2,    NA, 3, NA, NA),
                          c(NA, 1,    2,    NA, 3, NA, 4),
                          c(NA, NA, 5,    NA,6, NA, NA),
                          c(NA, NA, NA, NA,7, NA, NA),
                          c(15,    NA,    NA, NA,   NA, NA, NA),
                          c(NA, 9,    10,  NA,  11, NA, NA),
                          c(NA, 9,    10,  NA,  11, NA, NA),
                          c(NA, NA, 12,   NA, 13, NA, NA),
                          c(NA, NA, NA, NA,14, NA, NA)),
                      widths=c(0.7, 4, 0.7, 0.1, 5, 0.7, 3),
                      heights=rep(c(1, 1, 6, 1, 2), 2))

    ggsave(file = paste0("cache/alpha-beta-", key , ".png"), plot=g4, height=6, width=9)

    plotlist[[key]] = g4
}

g <- arrangeGrob(grobs = plotlist, ncol=2)
ggsave(file="cache/plotlist.png", g, height=8*length(plotlist)/1.5, width=10*2)
```

<img src="cache/plotlist.png"/>

# Alpha scatter

```{r draw alpha distribution scatter plot}
scatterplot.df <- lapply(ret, function(x) x$alphadistribution) %>%
    Reduce(function(x, y) merge(x, y, all=TRUE, by="Sample_ID"), .) %>%
    select(-Sample_ID) %>%
    na.omit()

galphascatter <- ggpairs(scatterplot.df,
                         lower = list(continuous = wrap("points",
                                                        alpha = 0.3,
                                                        size=0.1), 
                                      combo = wrap("dot",
                                                   alpha = 0.4,
                                                   size=0.2))) +
    theme(legend.position = "none", 
          panel.grid.major = element_blank(), 
          axis.ticks = element_blank(), 
          panel.border = element_rect(linetype = "dashed", colour = "black", fill = NA))

ggsave(file = paste0("cache/ggpairs-alphas.png"),
       plot= galphascatter,
       height=6,
       width=9,
       unit = "in",
       dpi = 300)
```

<img src="cache/ggpairs-alphas.png"/>

# Beta diversity

Beta diversity R2

```{r beta diversity r2}
betadf <- data.frame()
for (key in names(rds)) {
    moddf.full <- Reduce(function(...) merge(..., all=T), ret[[key]][["betadiversity"]]$full) %>%
        dplyr::filter(response == term) %>%
        mutate(model = key, covariates = "full") %>%
        select(term, model, covariates, R2)
    moddf.min <- Reduce(function(...) merge(..., all=T), ret[[key]][["betadiversity"]]$min) %>%
        dplyr::filter(response == term) %>%
        mutate(model = key, covariates = "min") %>%
        select(term, model, covariates, R2)
    betadf <- rbind(betadf, moddf.full, moddf.min)
}

betadf %>% arrange(term) %>% spread(model, R2) %>% kable
```

Beta diversity p.values

```{r beta diversity pvalue}
betadf <- data.frame()
for (key in names(rds)) {
    moddf.full <- Reduce(function(...) merge(..., all=T), ret[[key]][["betadiversity"]]$full) %>%
        dplyr::filter(response == term) %>%
        mutate(model = key, covariates = "full") %>%
        select(term, model, covariates, `Pr(>F)`)
    moddf.min <- Reduce(function(...) merge(..., all=T), ret[[key]][["betadiversity"]]$min) %>%
        dplyr::filter(response == term) %>%
        mutate(model = key, covariates = "min") %>%
        select(term, model, covariates, `Pr(>F)`)
    betadf <- rbind(betadf, moddf.full, moddf.min)
}
betadf %>% arrange(term) %>% spread(model, `Pr(>F)`) %>% kable
```

>>>>

for (key in names(rds)) {
    pseq <- import_filter_data(rds[[key]],
                               included = c("MEN", "BL_AGE", "SYSTM", "DIASM", "BMI",
                                            "CURR_SMOKE", "BL_USE_RX_C09",
                                            "Q57X", "BL_USE_RX_C03",
                                            "BL_USE_RX_C07", "BL_USE_RX_C08"))

    bray.dist.m.species <- calculate.beta.matrix(pseq)

    adonis.species <- calculate.betadiversity(pseq = pseq.species,
                                               matrix = bray.dist.m.species,
                                               vars = list("max" = var.CL,
                                                            "min" = var.CL.min))

    alphadiversity <- calculate.alphadiversity(pseq.species,
                                               list("max" = var.CL,
                                                    "min" = var.CL.min))

    alphadiversity <- calculate.alphadiversity(pseq.species,
                                               list("max" = var.CL,
                                                    "min" = var.CL.min))

    diversities <- mergediversities(alphadiversity, adonis.species, names.dset = names.dset)

    g4 <- plot.diversities(diversities)
    ggsave(file = sprintf("cache/comparepseqs-%s.png", key), plot=g4, height=6, width=9)
}
```
